---
title: "Applied Analytics Final Project"
author: "Zhongyi Zhang, Qi'er An, Yuying Gao"
date: "3/13/2019"
output: html_document
---

## R Markdown Data Cleaning:

```{r setup, include=FALSE}
rm(list = ls())
library(plyr)
library(dplyr)
library(lubridate)
library(ggplot2)
# "Region" CSV file data cleaning
d.in<-read.csv("~/Desktop/RegionIP2.csv")
#if(d.in$influenza.deaths == 0){d.in$influenza.deaths <- d.in$influenza.deaths + 1}
d.in2 <- d.in %>%
  filter(d.in$influenza.deaths == 0)
d.in <- filter(d.in, d.in$influenza.deaths != 0)
d.in2$influenza.deaths <- d.in2$influenza.deaths + 1
d.in <- rbind(d.in, d.in2)
d.in <- d.in %>%
  arrange(region, MMWR)
rm(data.frame = d.in2)
#d.in <- d.in%>%
#  -c(d.in$influenza.deaths == 0)
#d.in2$influenza.deaths <- d.in2$influenza.deaths + 1
#d.in$influenza.deaths <- d.in$influenza.deaths + 1
d.in <- d.in %>%
  mutate(ratio = pneumonia.deaths/influenza.deaths) 
#d.in$ratio[d.in$ratio == Inf] <- 1000
d.in <- d.in %>%
  mutate(year = floor(MMWR/100))
d.in <- d.in %>%
  mutate(week = MMWR - round(MMWR, -2))
d.in$week[d.in$week == -50] <- 50
d.in$week[d.in$week == -49] <- 51
d.in$week[d.in$week == -48] <- 52
d.in$week[d.in$week == -47] <- 53
d.in <- d.in %>%
  mutate(season = week)
#d.in <- data.frame(A= d.in$season(1:20),B=d.in$season(21:53))
#d.in$season[d.in$season[0:9]] <- "Winter"
df.1 <- subset(d.in, season>=9 & season<=21)
df.2 <- subset(d.in, season>=22 & season<=34)
df.3 <- subset(d.in, season>=35 & season<=47)
df.4 <- subset(d.in, season>=48 & season<=53)
df.5 <- subset(d.in, season>0 & season<=8)
df.1$season <- "1"
df.2$season <- "2"
df.3$season <- "3"
df.4$season <- "4"
df.5$season <- "4"
df.w <- rbind(df.5,df.4)
df.m <- rbind(df.1,df.2)
df.n <- rbind(df.3,df.w)
d.in0 <- rbind(df.m, df.n)
d.in0 <- d.in0 %>%
  arrange(region, MMWR)
rm(data.frame = d.in,df.1,df.2,df.3,df.4,df.5,df.m,df.n,df.w)

# "Age" CSV file data cleaning
d.out<-read.csv("~/Desktop/AgeIP2.csv",header=TRUE)
d.out2 <- d.out %>%
  filter(d.out$influenza.deaths ==0)
d.out <- filter(d.out, d.out$influenza.deaths != 0)
d.out2$influenza.deaths <- d.out2$influenza.deaths + 1
d.out <- rbind(d.out, d.out2)
d.out <- d.out %>%
  arrange(MMWR, age)
rm(data.frame = d.out2)
d.out <- d.out %>%
  mutate(agegroup = age)
d.out$agegroup <- as.character(d.out$agegroup)
d.out$agegroup[d.out$agegroup == "0-17 years"] <- 1
d.out$agegroup[d.out$agegroup == "18-64 years"] <- 2
d.out$agegroup[d.out$agegroup == "65 years and older"] <- 3
d.out <- d.out %>%
  mutate(ratio = pneumonia.deaths/influenza.deaths) 
d.out <- d.out %>%
  mutate(year = floor(MMWR/100))
d.out <- d.out %>%
  mutate(week = MMWR - round(MMWR, -2))
d.out$week[d.out$week == -50] <- 50
d.out$week[d.out$week == -49] <- 51
d.out$week[d.out$week == -48] <- 52
d.out$week[d.out$week == -47] <- 53
d.out <- d.out %>%
  mutate(season = week)
df.6 <- subset(d.out, season>=9 & season<=21)
df.7 <- subset(d.out, season>=22 & season<=34)
df.8 <- subset(d.out, season>=35 & season<=47)
df.9 <- subset(d.out, season>=48 & season<=53)
df.10 <- subset(d.out, season>0 & season<=8)
df.6$season <- "1"
df.7$season <- "2"
df.8$season <- "3"
df.9$season <- "4"
df.10$season <- "4"
df.o <- rbind(df.10,df.9)
df.p <- rbind(df.6,df.7)
df.q <- rbind(df.8,df.o)
d.out0 <- rbind(df.p, df.q)
d.out0 <- d.out0 %>%
  arrange(MMWR, agegroup)
rm(data.frame = d.out,df.6,df.7,df.8,df.9,df.10,df.o,df.p,df.q)

```




```{r}
###Table 1
d.in0 <- d.in0 %>%
  mutate(region = as.factor(region), season = as.factor(season),year = as.factor(year),pneumonia.deaths=as.numeric(pneumonia.deaths))

d.out0 <- d.out0 %>%
  mutate(agegroup = as.factor(agegroup), season = as.factor(season),year = as.factor(year))

intotalr = sum(d.in0$influenza.deaths)
intotalr
pntotalr = sum(d.in0$pneumonia.deaths)
pntotalr

intotala = sum(d.out0$influenza.deaths)
intotala
pntotala = sum(d.out0$pneumonia.deaths)
pntotala

#count region
d.table1region <- d.in0 %>%
  group_by(region) %>%
  summarize(infcount = sum(influenza.deaths),pncount = sum(pneumonia.deaths))%>%
  mutate(infperc = infcount/intotalr*100,pnperc = pncount/pntotalr*100)

#count season region chart
d.table1rseason <- d.in0 %>%
  group_by(season) %>%
  summarize(infcount = sum(influenza.deaths),pncount = sum(pneumonia.deaths))%>%
  mutate(infperc = infcount/intotalr*100,pnperc = pncount/pntotalr*100)

#count year region chart
d.table1ryear <- d.in0 %>%
  group_by(year) %>%
  summarize(infcount = sum(influenza.deaths),pncount = sum(pneumonia.deaths))%>%
  mutate(infperc = infcount/intotalr*100,pnperc = pncount/pntotalr*100)


#count age
d.table1age <- d.out0 %>%
  group_by(agegroup) %>%
  summarize(infcount = sum(influenza.deaths),pncount = sum(pneumonia.deaths))%>%
  mutate(infperc = infcount/intotala*100,pnperc = pncount/pntotala*100)

#count season age chart
d.table1aseason <- d.out0 %>%
  group_by(season) %>%
  summarize(infcount = sum(influenza.deaths),pncount = sum(pneumonia.deaths))%>%
  mutate(infperc = infcount/intotala*100,pnperc = pncount/pntotala*100)

#count year age chart
d.table1ayear <- d.out0 %>%
  group_by(year) %>%
  summarize(infcount = sum(influenza.deaths),pncount = sum(pneumonia.deaths))%>%
  mutate(infperc = infcount/intotalr*100,pnperc = pncount/pntotalr*100)


```

```{r}
d.rregion1 <- d.in0 %>%
  filter(d.in0$region == 1)
d.rregion2 <- d.in0 %>%
  filter(d.in0$region == 2)
d.rregion3 <- d.in0 %>%
  filter(d.in0$region == 3)
d.rregion4 <- d.in0 %>%
  filter(d.in0$region == 4)
d.rregion5 <- d.in0 %>%
  filter(d.in0$region == 5)
d.rregion6 <- d.in0 %>%
  filter(d.in0$region == 6)
d.rregion7 <- d.in0 %>%
  filter(d.in0$region == 7)
d.rregion8 <- d.in0 %>%
  filter(d.in0$region == 8)
d.rregion9 <- d.in0 %>%
  filter(d.in0$region == 9)
d.rregion10 <- d.in0 %>%
  filter(d.in0$region == 10)

plot(x = d.in0$influenza.deaths, y = d.in0$pneumonia.deaths, col = d.in0$region, pch = 20)
abline(lm(d.rregion1$pneumonia.deaths ~ d.rregion1$influenza.deaths))
abline(lm(d.rregion2$pneumonia.deaths ~ d.rregion2$influenza.deaths))
abline(lm(d.rregion3$pneumonia.deaths ~ d.rregion3$influenza.deaths))
abline(lm(d.rregion4$pneumonia.deaths ~ d.rregion4$influenza.deaths))
abline(lm(d.rregion5$pneumonia.deaths ~ d.rregion5$influenza.deaths))
abline(lm(d.rregion6$pneumonia.deaths ~ d.rregion6$influenza.deaths))
abline(lm(d.rregion7$pneumonia.deaths ~ d.rregion7$influenza.deaths))
abline(lm(d.rregion8$pneumonia.deaths ~ d.rregion8$influenza.deaths))
abline(lm(d.rregion9$pneumonia.deaths ~ d.rregion9$influenza.deaths))
abline(lm(d.rregion10$pneumonia.deaths ~ d.rregion10$influenza.deaths))
```

```{r}
ggplot(d.in0,aes(influenza.deaths,y = pneumonia.deaths, group = region,col = region)) + geom_point() +labs(x = "death from influenza",y = "death from pneumonia")

ggplot(d.out0,aes(influenza.deaths,y = pneumonia.deaths, group = agegroup,col = agegroup)) + geom_point() +labs(x = "death from influenza",y = "death from pneumonia")

ggplot(d.in0,aes(influenza.deaths,y = pneumonia.deaths, group = season,col = season)) + geom_point() +labs(x = "death from influenza",y = "death from pneumonia")

ggplot(d.out0,aes(influenza.deaths,y = pneumonia.deaths, group = season,col = season)) + geom_point() +labs(x = "death from influenza",y = "death from pneumonia")

ggplot(d.in0,aes(influenza.deaths,y = pneumonia.deaths, group = year,col = year)) + geom_point() +labs(x = "death from influenza",y = "death from pneumonia")

ggplot(d.out0,aes(influenza.deaths,y = pneumonia.deaths, group = year,col = year)) + geom_point() +labs(x = "death from influenza",y = "death from pneumonia")
```

```{r}
#Examples for single regression
m.rr <- lm(ratio ~ region,data = d.in0)
summary(m.rr)
plot(m.rr)

pv.rr <- summary(m.rr)$coefficient[,"Pr(>|t|)"]
names(pv.rr) <- NULL
pv.rr

ggplot(data = d.in0, aes(x = region, y = ratio))+geom_point(color='blue')+geom_smooth(color='red',data = d.in0, aes(x=region, y=ratio))
ggplot(data = d.in0, aes(x = region, y = ratio))+geom_point(color='blue')+geom_smooth(color='red',method = 'lm',data = d.in0, aes(x=region, y=ratio))
eqn.rr <-paste0("ratio = ",round(m.rr$coefficients["region"],3),"region+",round(m.rr$coefficients["(Intercept)"],3),"; rsq = ",round(summary(m.rr)$r.squared,3))
ggplot(data.frame(d.in0$region,d.in0$ratio), aes(x=d.in0$region, y=d.in0$ratio))+geom_point()+ stat_smooth(method=lm) + annotate("text", label=eqn.rr, parse=F,x=8.3,y=750)
# Since rsq is only 0.016, ratio and region don't show a strong linear relationship

```
Interpretation:
In simple regression between ratio and region,
H0: beta = 0, 
HA: beta != 0
The p-value for ratio and region is
6.494317e-11
The p-value for ratio and region is far below 0.05. We need to reject the null hypothesis for ratio and region. Region have a statistically significant relationship with ratio.


```{r}
m.rs <- lm(ratio ~ season,data = d.in0)
summary(m.rs)
plot(m.rs)

pv.rs <- summary(m.rs)$coefficient[,"Pr(>|t|)"]
names(pv.rs) <- NULL
pv.rs

ggplot(data = d.in0, aes(x = season, y = ratio))+geom_point(color='blue')+geom_smooth(color='red',data = d.in0, aes(x=season, y=ratio))
ggplot(data = d.in0, aes(x = season, y = ratio))+geom_point(color='blue')+geom_smooth(color='red',method = 'lm',data = d.in0, aes(x=season, y=ratio))
eqn.rs <-paste0("ratio = ",round(m.rs$coefficients["season"],3),"season+",round(m.rs$coefficients["(Intercept)"],3),"; rsq = ",round(summary(m.rs)$r.squared,3))
ggplot(data.frame(d.in0$season,d.in0$ratio), aes(x=d.in0$season, y=d.in0$ratio)) + geom_point()+stat_smooth(method=lm) + annotate("text", label=eqn.rs, parse=F,x=3.75,y=750)
# Since rsq is 0.329, ratio and season somewhat show the linear relationship

```

Interpretation:
In simple regression between ratio and seaon,
H0: beta = 0, 
HA: beta != 0
The p-values for ratio vs. season2 (Summer), season3 (Autumn), and season4 (Winter) are 
1.764008e-133  
2.343433e-97
0.03465084e
The p-values for ratio vs. Summer and Autumn are far below 0.05. We need to reject the null hypothesis for ratio versus Summer and Autumn. Summer and Autumn have a statistically significant relationship with ratio.

The p-value for ratio vs. Winter is a little bit below 0.05. We somewhat need to reject the null hypothesis for ratio versus Winter. Winter have a some relationship with ratio.



```{r}
m.wi <- lm(d.out0$influenza.deaths ~ d.out0$week)
summary(m.wi)
confint(m.wi)
```
```{r}
m.wp <- lm(d.out0$pneumonia.deaths ~ d.out0$week)
summary(m.wp)
confint(m.wp)
```

```{r}
m.wi1 <- lm(d.in0$influenza.deaths ~ d.in0$week)
summary(m.wi1)
confint(m.wi1)
```

```{r}
m.wp2 <- lm(d.in0$pneumonia.deaths ~ d.in0$week)
summary(m.wp2)
confint(m.wp2)
```


```{r}
m.lmregion <- lm(pneumonia.deaths ~ influenza.deaths + region + season + year + week, data = d.in0)
summary(m.lmregion)
```

```{r}
m.lmage <- lm(pneumonia.deaths ~ influenza.deaths + agegroup + season + year + week, data = d.out0)
summary(m.lmage)
```



## R Markdown Data Analysis for Region CSV file

```{r}
m.r <- lm(ratio ~ region + season + year + week,data = d.in0)
summary(m.r)
plot(m.r)

pv.r <- summary(m.r)$coefficient[,"Pr(>|t|)"]
names(pv.r) <- NULL
pv.r

ggplot(data = d.in0, aes(x = region, y = ratio))+geom_point(color='blue')+geom_smooth(color='red',data = d.in0, aes(x=region, y=ratio))
ggplot(data = d.in0, aes(x = region, y = ratio))+geom_point(color='blue')+geom_smooth(color='red',method = 'lm',data = d.in0, aes(x=region, y=ratio))
eqn.r <-paste0("ratio = ",round(m.r$coefficients["region"],3),"region+",round(m.r$coefficients["(Intercept)"],3),"; rsq = ",round(summary(m.r)$r.squared,3))
ggplot(data.frame(d.in0$region,d.in0$ratio), aes(x=d.in0$region, y=d.in0$ratio))+geom_point()+ stat_smooth(method=lm) + annotate("text", label=eqn.r, parse=F,x=8.3,y=750)
# Since rsq is only 0.016, ratio and region don't show a strong linear relationship

```
Interpretation:
In simple regression between ratio and region,
H0: beta = 0, 
HA: beta != 0
The p-value for ratio and region is
6.494317e-11
The p-value for ratio and region is far below 0.05. We need to reject the null hypothesis for ratio and region. Region have a statistically significant relationship with ratio.

```{r}
m.a <- lm(ratio ~ agegroup + season + year + week,data = d.out0)
summary(m.a)
plot(m.a)

pv.a <- summary(m.a)$coefficient[,"Pr(>|t|)"]
names(pv.a) <- NULL
pv.a

ggplot(data = d.out0, aes(x = agegroup, y = ratio))+geom_point(color='blue')+geom_smooth(color='red',data = d.out0, aes(x=agegroup, y=ratio))
ggplot(data = d.out0, aes(x = agegroup, y = ratio))+geom_point(color='blue')+geom_smooth(color='red',method = 'lm',data = d.out0, aes(x=agegroup, y=ratio))
eqn.a <-paste0("ratio = ",round(m.a$coefficients["agegroup"],3),"agegroup+",round(m.a$coefficients["(Intercept)"],3),"; rsq = ",round(summary(m.a)$r.squared,3))
ggplot(data.frame(d.out0$agegroup,d.out0$ratio), aes(x=d.out0$agegroup, y=d.out0$ratio))+geom_point()+ stat_smooth(method=lm) + annotate("text", label=eqn.a, parse=F,x=8.3,y=750)
```




Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

```{r}
kmeans_in <- kmeans(d.in0,centers=2)
names(kmeans_in)
kmeans_in$cluster
kmeans_in$centers
plot(d.in0$week, d.in0$ratio,col=kmeans_in$cluster,xlab = "week", ylab ="ratio",main = "k-means plot for $k$ = 2, region chart",pch=20,cex=2)
points(kmeans_in$centers,col=1:2,pch=3,cex=3,lwd=3)

KMeans_in <- kmeans(d.in0,centers=2)
plot(d.in0$week, d.in0$ratio,col=kmeans_in$cluster,xlab = "week", ylab ="ratio",main = "k-means plot for $k$ = 3",pch=19,cex=2)
points(kmeans_in$centers,pch=3,cex=4,lwd=3, col="purple")
print(kmeans_in$betweenss)
print(kmeans_in$tot.withinss)
```

```{r}
kmeans_in <- kmeans(d.in0,centers=3)
names(kmeans_in)
kmeans_in$cluster
kmeans_in$centers
plot(d.in0$week, d.in0$ratio,col=kmeans_in$cluster,xlab = "week", ylab ="ratio",main = "k-means plot for $k$ = 3, region chart",pch=20,cex=2)
points(kmeans_in$centers,col=1:3,pch=3,cex=3,lwd=3)

KMeans_in <- kmeans(d.in0,centers=3)
plot(d.in0$week, d.in0$ratio,col=kmeans_in$cluster,xlab = "week", ylab ="ratio",main = "k-means plot for $k$ = 3",pch=20,cex=2)
points(kmeans_in$centers,pch=3,cex=4,lwd=3, col="purple")
print(kmeans_in$betweenss)
print(kmeans_in$tot.withinss)
```


```{r}
kmeans_in <- kmeans(d.in0,centers=4)
names(kmeans_in)
kmeans_in$cluster
kmeans_in$centers
plot(d.in0$week, d.in0$ratio,col=kmeans_in$cluster,xlab = "week", ylab ="ratio",main = "k-means plot for $k$ = 4, region chart",pch=20,cex=2)
points(kmeans_in$centers,col=1:4,pch=3,cex=3,lwd=3)

KMeans_in <- kmeans(d.in0,centers=4)
plot(d.in0$week, d.in0$ratio,col=kmeans_in$cluster,xlab = "week", ylab ="ratio",main = "k-means plot for $k$ = 3",pch=20,cex=2)
points(kmeans_in$centers,pch=3,cex=4,lwd=3, col="purple")
print(kmeans_in$betweenss)
print(kmeans_in$tot.withinss)

```

```{r}
kmeans_in <- kmeans(d.in0,centers=5)
names(kmeans_in)
kmeans_in$cluster
kmeans_in$centers
plot(d.in0$week, d.in0$ratio,col=kmeans_in$cluster,xlab = "week", ylab ="ratio",main = "k-means plot for $k$ = 5, region chart",pch=20,cex=2)
points(kmeans_in$centers,col=1:5,pch=3,cex=3,lwd=3)

KMeans_in <- kmeans(d.in0,centers=5)
plot(d.in0$week, d.in0$ratio,col=kmeans_in$cluster,xlab = "week", ylab ="ratio",main = "k-means plot for $k$ = 3",pch=20,cex=2)
points(kmeans_in$centers,pch=3,cex=4,lwd=3, col="purple")
print(kmeans_in$betweenss)
print(kmeans_in$tot.withinss)
```
```{r}
kmeans_in <- kmeans(d.in0,centers=6)
names(kmeans_in)
kmeans_in$cluster
kmeans_in$centers
plot(d.in0$week, d.in0$ratio,col=kmeans_in$cluster,xlab = "week", ylab ="ratio",main = "k-means plot for $k$ = 6, region chart",pch=20,cex=2)
points(kmeans_in$centers,col=1:6,pch=3,cex=3,lwd=3)

KMeans_in <- kmeans(d.in0,centers=6)
plot(d.in0$week, d.in0$ratio,col=kmeans_in$cluster,xlab = "week", ylab ="ratio",main = "k-means plot for $k$ = 3",pch=20,cex=2)
points(kmeans_in$centers,pch=3,cex=4,lwd=3, col="purple")
print(kmeans_in$betweenss)
print(kmeans_in$tot.withinss)
```



```{r}
d.ou0 <- subset(d.out0,select = -c(age,MMWR,influenza.deaths,pneumonia.deaths))

kmeans_ou <- kmeans(d.ou0,centers=2)

names(kmeans_ou)
kmeans_ou$cluster
kmeans_ou$centers

plot(d.ou0$week, d.ou0$ratio,col=kmeans_ou$cluster,xlab = "week", ylab ="ratio",pch=20,cex=2,main = "k-means plot for $k$ = 2, age chart")
points(kmeans_ou$centers,col=1:2,pch=3,cex=3,lwd=3)
print(kmeans_ou$betweenss)
print(kmeans_ou$tot.withinss)

```


```{r}
kmeans_ou <- kmeans(d.ou0,centers=3)

names(kmeans_ou)
kmeans_ou$cluster
kmeans_ou$centers

plot(d.ou0$week, d.ou0$ratio,col=kmeans_ou$cluster,xlab = "week", ylab ="ratio",pch=20,cex=2,main = "k-means plot for $k$ = 3, age chart")
points(kmeans_ou$centers,col=1:3,pch=3,cex=3,lwd=3)
print(kmeans_ou$betweenss)
print(kmeans_ou$tot.withinss)

```

```{r}
kmeans_ou <- kmeans(d.ou0,centers=4)

names(kmeans_ou)
kmeans_ou$cluster
kmeans_ou$centers

plot(d.ou0$week, d.ou0$ratio,col=kmeans_ou$cluster,xlab = "week", ylab ="ratio",pch=20,cex=2,main = "k-means plot for $k$ = 4, age chart")
points(kmeans_ou$centers,col=1:4,pch=3,cex=3,lwd=3)
print(kmeans_ou$betweenss)
print(kmeans_ou$tot.withinss)

```

```{r}
kmeans_ou <- kmeans(d.ou0,centers=5)

names(kmeans_ou)
kmeans_ou$cluster
kmeans_ou$centers

plot(d.ou0$week, d.ou0$ratio,col=kmeans_ou$cluster,xlab = "week", ylab ="ratio",pch=20,cex=2,main = "k-means plot for $k$ = 5, age chart")
points(kmeans_ou$centers,col=1:5,pch=3,cex=3,lwd=3)
print(kmeans_ou$betweenss)
print(kmeans_ou$tot.withinss)

```

```{r}
kmeans_ou <- kmeans(d.ou0,centers=6)

names(kmeans_ou)
kmeans_ou$cluster
kmeans_ou$centers

plot(d.ou0$week, d.ou0$ratio,col=kmeans_ou$cluster,xlab = "week", ylab ="ratio",pch=20,cex=2,main = "k-means plot for $k$ = 6, age chart")
points(kmeans_ou$centers,col=1:6,pch=3,cex=3,lwd=3)
print(kmeans_ou$betweenss)
print(kmeans_ou$tot.withinss)

```